Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-09-10 ms.tex
Page 1 edits finished

%%% FOR DRAFTING: REMOVE AFTER PAPER IS COMPLETE!
\usepackage{xcolor}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\proposed}[1]{\textcolor{blue}{#1}}

@@

The outer layers of planetary atmospheres are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation and regulate the escape of gas into space. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in this process as the optical depth at line center is large. Ionization and subsequent recombination can generate \lya within the planet's atmosphere, dissociating molecules in lower layers and causing heating by collisional de-excitation. \lya can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy.

Hubble Space Telescope (HST) observations with STIS have found large \lya transit depths around a handful of exoplanets, indicating an extended distribution of hydrogen: \citet{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order the planetary radius for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with H atoms surrounding a planet. The importance of the \lya line in the atmospheric dynamics of these systems necessitates a careful treatment of resonant scattering in order to construct accurate models of atmospheric heating and escape. However, the extreme optical depths at \lya line center introduce a computational barrier to simulating radiative transfer via Monte Carlo methods.

>>>>

@@

The outer layers of planetary atmospheres are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation as well as regulate the escape of gas into space. Ionization and subsequent recombination can generate \lya within the planet's atmosphere, dissociating molecules in lower layers and creating pressure forces that drive an outflow of escaping H atoms. \lya can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy \citep{2017ApJ...851..150H}.

Hubble Space Telescope (HST) observations with STIS have found large \lya transit depths around a handful of exoplanets \citep{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order a few planetary radii for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with the H atom population. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in the atmospheric dynamics of these systems as the optical depth at line center is large. The importance of the \lya line in this process necessitates a careful treatment of resonant scattering in order to construct accurate models of H atom excitation, heating, and radiative forces. However, the extreme optical depths at \lya line center introduce a computational barrier to simulating radiative transfer via Monte Carlo methods.
